\section{Organization of the proof files}
\label{appendix:proof-folders}

\subheading{Requirements.}
Our proofs requires the use of \emph{Coq 8.8.2} for the proofs and
\emph{Opam 2.0} to manage the dependencies. We are aware that there exists more
recent versions of Coq; VST; CompCert etc. however to avoid dealing with backward
breaking compatibility we decided to freeze our dependencies.

\subheading{Associated files.}
The repository containing the proof is composed of two folders \textbf{\texttt{packages}}
and \textbf{\texttt{proofs}}.
It aims to be used at the same time as an \emph{opam} repository to manage
the dependencies of the proof and to provide the code.

The actual proofs can be found in the \textbf{\texttt{proofs}} folder in which
the reader will find the directories \textbf{\texttt{spec}} and \textbf{\texttt{vst}}.

\subheading{\texttt{packages/}}
This folder makes sure that we are using the correct version of
Verifiable Software Toolchain (version 2.0) and CompCert (version 3.2).
Additionally it pins the version of the elliptic curves library by Bartzia and Strub
and allows us to use the theorem of quadratic reciprocity.

\subheading{\texttt{proofs/spec/}}
In this folder the reader will find multiple levels of implementation of X25519.
\begin{itemize}
  \item \textbf{\texttt{Libs/}} contains basic libraries and tools to help use
        reason with lists and decidable procedures.
  \item \textbf{\texttt{ListsOp/}} defines operators on list such as
        \Coqe{ZofList} and related lemmas using \eg \VSTe{Forall}.
  \item \textbf{\texttt{Gen/}} defines a generic Montgomery ladder which can be
        instantiated with different operations. This ladder is the stub for the
        following implementations.
  \item \textbf{\texttt{High/}} contains the theory of Montgomery curves,
        twists, quadratic extensions and ladder.
        It also proves the correctness of the ladder over $\F{\p}$.
  \item \textbf{\texttt{Mid/}} provides a list-based implementation of the
        basic operations \TNaCle{A}, \TNaCle{Z}, \TNaCle{M} \ldots~and the ladder. It
        makes the link with the theory of Montgomery curves.
  \item \textbf{\texttt{Low/}} provides a second list-based implementation of
        the basic operations \TNaCle{A}, \TNaCle{Z}, \TNaCle{M} \ldots~and the ladder.
        Those functions are proven to provide the same results as the ones in
        \texttt{Mid/}, however their implementation are closer to \texttt{C} in order
        facilitate the proof of equivalence with TweetNaCl code.
  \item \textbf{\texttt{rfc/}} provides our rfc formalization.
        It uses integers for the basic operations \TNaCle{A}, \TNaCle{Z}, \TNaCle{M}
        \ldots and the ladder. It specifies the decoding/encoding of/to byte
        arrays (seen as list of integers) as in RFC~7748.
\end{itemize}

\subheading{\texttt{proofs/vst/}}
Here the reader will find four folders.
\begin{itemize}
  \item \textbf{\texttt{c}} contains the C Verifiable implementation of TweetNaCl.
        \texttt{clightgen} will generate the appropriate translation into Clight.
  \item \textbf{\texttt{init}} contains basic lemmas and memory manipulation
        shortcuts to handle the aliasing cases.
  \item \textbf{\texttt{spec}} defines as Hoare triple the specification of the
        functions used in \TNaCle{crypto_scalarmult}.
  \item \textbf{\texttt{proofs}} contains the proofs of the above Hoare triples
        and thus the proof that TweetNaCl code is sound and correct.
\end{itemize}
